---
title: "Group Assignment Training SME"
author: "Nzambuli Daniel"
date: "2023-11-15"
output: html_document
---

```{r}
library(readxl)
SME_Training_Data <- read_excel("SME Training Data.xlsx")
head(SME_Training_Data)
```

# Based on Region

**Task**

1.  Comparing diff regions to see what ways they are similar & where they differ
2.  The `response variable` if they need the facility for training

***Note:*** Include binary & ordinal responses but disregard categorical

```{r}
str(SME_Training_Data)
```

**Ordinal Data**

1.  Age group of Respondents `Age`
2.  Education level `Education`
3.  Region where the SME is located `Location`
4.  Business size(number of employees) `Size`
5.  Age of the business `Ageofbusiness`

**Binary Data**

1.  Gender of respondent `Gender`
2.  Facility needed `Facility`
3.  Presence of Disabled persons `DAP`

## Ruaraka

Facility `2` is No and `1` is yes encoding is done such that yes becomes `1` and no becomes `0`

```{r}
library(dplyr, warn.conflicts = FALSE)

SME_Training_Data = SME_Training_Data %>% mutate(Facility = case_when(
  Facility == 1 ~ 1,
  Facility == 2 ~ 0,
  TRUE ~ Facility
))
SME_Training_Data$Location[is.na(SME_Training_Data$Location)] = 10
```

Ruaraka is location `1` this means all other locations are `0`

```{r}
ruaraka_data = SME_Training_Data %>% mutate(Location = case_when(
  Location == 1~1,
  Location == 2~1,
  Location == 7~1,
  Location == 9~1,
  Location == 10 ~ 1,
  TRUE ~Location
))
ruaraka_data = ruaraka_data %>% mutate(Location = case_when(
  Location != 1~0,
  TRUE ~Location
))
```

### Drop all columns that are not Binary or Ordinal

```{r}
colnames(SME_Training_Data)
```

**Key columns**

1.  Age - 3
2.  Education - 4
3.  Location - 6
4.  Size - 7
5.  AgeofBusiness - 8
6.  DAP - 9
7.  Gender = 2
8.  Facility = 12

```{r}
ruaraka_data = ruaraka_data[, c(2, 3, 4, 6, 7, 8, 9, 12)]
colnames(ruaraka_data)
```

```{r}
ruaraka_cols = colnames(ruaraka_data)[1: 7]
ruaraka_cols
```

### contingency tables

```{r}
for(col in ruaraka_cols){
  print(col)
  print(table(facility = ruaraka_data$Facility,
              ruaraka_data[[col]]))
}
```

**Remove all values less than 5 to perform a chi-square test**

```{r}
for(col in ruaraka_cols){
  print(col)
  print(table(facility = ruaraka_data$Facility,
              ruaraka_data[[col]])[rowMeans(table(facility = ruaraka_data$Facility,
              ruaraka_data[[col]])) >= 5,
              colMeans(table(facility = ruaraka_data$Facility,
              ruaraka_data[[col]])) >= 5])
  cat("\n\n\n")
  print(paste("Chi-squred test of", col))
  print(suppressWarnings(chisq.test(table(facility = ruaraka_data$Facility,
              ruaraka_data[[col]])[rowMeans(table(facility = ruaraka_data$Facility,
              ruaraka_data[[col]])) >= 5,
              colMeans(table(facility = ruaraka_data$Facility,
              ruaraka_data[[col]])) >= 5])))
}
```

**Interpret**

`Location` is the only data that can significantly predict whether an individual needs facilities in **Ruaraka**

### Logistic model

```{r}
library(caTools)
library(ROCR)
set.seed(123)
split_ruaraka =  sample(1:nrow(ruaraka_data), 0.8* nrow(ruaraka_data))
train_ruaraka = ruaraka_data[split_ruaraka,]
test_ruaraka = ruaraka_data[-split_ruaraka,]
```

```{r}
cat("Training data has", nrow(train_ruaraka), "rows\n")
cat("Test data has", nrow(test_ruaraka), "rows\n")
```

```{r}
ruaraka_model = glm(Facility~Location,
                    data = train_ruaraka, family = binomial)
summary(ruaraka_model)
```

### Prediction and Accuracy

```{r}
pred_ruaraka = predict(ruaraka_model, 
                       test_ruaraka,
                       type = "response")
pred_ruaraka = ifelse(pred_ruaraka> 0.5, 1, 0)
reg_tab_ruaraka = table(predicted_facility = test_ruaraka$Facility, pred_ruaraka)
reg_tab_ruaraka
```

```{r}
odds_rat_ruaraka = exp(coef(ruaraka_model))
odds_rat_ruaraka
```

This means that if a person is from `ruaraka` they will need the facilities in the university

$$
Need(faclity) = 2.088235 + Ruaraka 0.000007
$$

so `0.0007%` of people from Ruaraka need the facilities with atleast `2` people who will need the facilities

## Embakasi

```{r}
embakasi_data = SME_Training_Data %>% mutate(Location = case_when(
  Location != 2~0,
  TRUE ~Location
))
embakasi_data = embakasi_data[, c(2, 3, 4, 6, 7, 8, 9, 12)]
colnames(embakasi_data)
```

```{r}
for(col in colnames(embakasi_data)[1: 7]){
  print(col)
  print(table(facility = embakasi_data$Facility,
              embakasi_data[[col]]))
}
```

```{r}
embakasi_cols = colnames(embakasi_data[1:7])
for(col in embakasi_cols){
  print(col)
  print(table(facility = embakasi_data$Facility,
              embakasi_data[[col]])[rowMeans(table(facility = embakasi_data$Facility,
              embakasi_data[[col]])) >= 5,
              colMeans(table(facility = embakasi_data$Facility,
              embakasi_data[[col]])) >= 5])
  cat("\n\n\n")
  print(paste("Chi-squred test of", col))
  print(suppressWarnings(chisq.test(table(facility = embakasi_data$Facility,
              embakasi_data[[col]])[rowMeans(table(facility = embakasi_data$Facility,
              embakasi_data[[col]])) >= 5,
              colMeans(table(facility = embakasi_data$Facility,
              embakasi_data[[col]])) >= 5])))
}
```

`Age of the Business` is the only data that can significantly predict whether an individual needs facilities in **Ruaraka**

```{r}
split_embakasi =  sample(1:nrow(embakasi_data), 0.8* nrow(embakasi_data))
train_embakasi = embakasi_data[split_embakasi,]
test_embakasi = embakasi_data[-split_embakasi,]
```

```{r}
embakasi_model = glm(Facility~Location,
                    data = train_embakasi, family = binomial)
summary(embakasi_model)
```
