---
title: "Group Assignment Training SME"
author: "Nzambuli Daniel"
date: "2023-11-15"
output: html_document
---

```{r}
library(readxl)
SME_Training_Data <- read_excel("SME Training Data.xlsx")
head(SME_Training_Data)
```

# Based on Region

## Add a columns for virtual learning, physical learning and combined learning 

The columns shall be named

1.  `face-2face` for physical
2.  `online_learn` for virtual
3.  `physical_virtual` for combined virtual and physical

```{r}
SME_Training_Data$face_2_face = as.factor(ifelse(SME_Training_Data$Learning == 1, 1, 0))
SME_Training_Data$online_learn = as.factor(ifelse(SME_Training_Data$Learning == 2, 1, 0))
SME_Training_Data$physical_virtual = as.factor(ifelse(SME_Training_Data$Learning == 3, 1, 0))
```

**Task**

1.  Comparing diff regions to see what ways they are similar & where they differ
2.  The `response variable` if they need the facility for training

***Note:*** Include binary & ordinal responses but disregard categorical

### Re-label the factor columns

## Exploritory analysis

### Type of data in each column

```{r}
str(SME_Training_Data)
```

### 

```{r}
# gender 2 female 1 male
SME_Training_Data$Gender = as.factor(ifelse(SME_Training_Data$Gender == 2, "Female", "MALE"))
```

```{r}
SME_Training_Data$Education = as.factor(ifelse(
  SME_Training_Data$Education == 1, 'Primary', ifelse(
    SME_Training_Data$Education == 2, 'Secondary',
    ifelse(SME_Training_Data$Education == 3, 'Diploma', 
           ifelse(
             SME_Training_Data$Education == 4, 'Undergraduate',
             ifelse(SME_Training_Data$Education == 5, 'Post Grad',
                    ifelse(SME_Training_Data$Education == 6, 'Other', SME_Training_Data$Education))
           ))
  )
))
```

```{r}
SME_Training_Data$Age = as.factor(ifelse(SME_Training_Data$Age == 1, 'below 20',
                                         ifelse(SME_Training_Data$Age == 2, '21 to 30',
                                                ifelse(SME_Training_Data$Age == 3, '31 to 40',ifelse(SME_Training_Data$Age == 4, '41 to 50',ifelse(SME_Training_Data$Age == 5, '51 to 60',ifelse(SME_Training_Data$Age == 6, 'above 60', SME_Training_Data$Age)))))))
```

```{r}
SME_Training_Data$Size = as.factor(ifelse(SME_Training_Data$Size == 1, '1', ifelse(
  SME_Training_Data$Size == 2, '2 to 4', ifelse(
    SME_Training_Data$Size == 3, '5 to 10',ifelse(
      SME_Training_Data$Size == 4, '11 to 49', ifelse(
        SME_Training_Data$Size == 5, 'above 50',SME_Training_Data$Size
      )
    )
  )
)))
```

```{r}
SME_Training_Data$Ageofbusiness = as.factor(ifelse(
  SME_Training_Data$Ageofbusiness == 1, '< 1 year',ifelse(
    SME_Training_Data$Ageofbusiness == 2, '1 to 3', ifelse(
      SME_Training_Data$Ageofbusiness == 3, '4 to 6', ifelse(
        SME_Training_Data$Ageofbusiness == 4, '6 to 10', ifelse(
          SME_Training_Data$Ageofbusiness == 5, '< 10 years', SME_Training_Data$Ageofbusiness
        )
      )
    )
  )
))
```

```{r}
SME_Training_Data$DAP = as.factor(ifelse(
  SME_Training_Data$DAP == 1, 'YES',ifelse(
    SME_Training_Data$DAP == 2, 'No', SME_Training_Data$DAP
  )
))
```

### Missing-ness of the data

```{r}
library(dplyr)
library(tidyverse)
library(finalfit)
```

```{r}
missing_plot(SME_Training_Data)
```

### **Figure out the number of data points for each location in the data**

```{r}
location = SME_Training_Data$Location
min(location, na.rm = TRUE)
max(location, na.rm = TRUE)
```

#### **Convert Na values to region 10**

```{r}
SME_Training_Data$Location[is.na(SME_Training_Data$Location)] = 10
```

```{r}
location = SME_Training_Data$Location
min(location)
max(location)
```

### Re-code the data back to string

```{r}
locality = data.frame(town = location, count = rep(1, length(location)))
locality = locality %>% group_by(town) %>% summarise(count = sum(count))
locality
```

```{r}
#  'Ruaraka' = 1
#  'Embakasi' = 2
#  'Langata' = 3
#  'Starehe' = 4
#  'Mathare' = 5
#  'Kibra' = 6
#  'Dagoretti' = 7
#  'Westlands' = 8
#  'Roysambu' = 9
#  'Kasarani' = 10
#  'Kamukunji' = 11
#  'Other' = 12
locality$town = ifelse(locality$town == 1, 'Ruaraka', ifelse(
  locality$town == 2, 'Embakasi',ifelse(
    locality$town == 3, 'Langata',ifelse(
      locality$town == 4, 'Starehe',ifelse(
        locality$town == 5, 'Mathare',ifelse(
          locality$town == 6, 'Kibra',ifelse(
            locality$town == 7, 'Dagoreti',ifelse(
              locality$town == 8, 'Westlands',ifelse(
                locality$town == 9, 'Roysambu',ifelse(
                  locality$town == 10, 'other',locality$town
                  )
                )
              )
            )
          )
        )
      )
    )
  )
  )
locality
```

**Ordinal Data**

1.  Age group of Respondents `Age`
2.  Education level `Education`
3.  Region where the SME is located `Location`
4.  Business size(number of employees) `Size`
5.  Age of the business `Ageofbusiness`

**Binary Data**

1.  Gender of respondent `Gender`
2.  Facility needed `Facility`
3.  Presence of Disabled persons `DAP`

The data for `location` was grouped into `near Nairobi` and `far from Nairobi` based on ease of walking from the town to the CBD

## Far from Nairobi

Far from towns are `Ruaraka, Embakasi, Dagoretti, Roysambu, Other`

Facility `2` is No and `1` is yes encoding is done such that yes becomes `1` and no becomes `0`

```{r}
library(dplyr, warn.conflicts = FALSE)

SME_Training_Data = SME_Training_Data %>% mutate(Facility = case_when(
  Facility == 1 ~ 1,
  Facility == 2 ~ 0,
  TRUE ~ Facility
))
SME_Training_Data$Location[is.na(SME_Training_Data$Location)] = 10
```

Far from Nairobi is location `1` this means all other locations are `0`

```{r}
far_data = SME_Training_Data %>% mutate(Location = case_when(
  Location == 1~1,
  Location == 2~1,
  Location == 7~1,
  Location == 9~1,
  Location == 10 ~ 1,
  TRUE ~Location
))
far_data = far_data %>% mutate(Location = case_when(
  Location != 1~0,
  TRUE ~Location
))
```

### Drop all columns that are not Binary or Ordinal

```{r}
colnames(SME_Training_Data)
```

**Key columns**

1.  Age - 3
2.  Education - 4
3.  Location - 6
4.  Size - 7
5.  AgeofBusiness - 8
6.  DAP - 9
7.  Gender = 2
8.  Facility = 12
9.  face_2_face = 17
10. online_learn = 18
11. physical_virtual = 19

```{r}
far_data = far_data[, c(2, 3, 4, 6, 7, 8, 9, 12, 17, 18, 19)]
colnames(far_data)
```

```{r}
far_cols = cbind(colnames(far_data)[1: 7], colnames(far_data)[9:11])
far_cols
```

### contingency tables

facility `1` is yes they need the facility and `0` is no they do not need it

```{r}
for(col in far_cols){
  print(col)
  print(table(facility = far_data$Facility,
              far_data[[col]]))
}
```

**Remove all values less than 5 to perform a chi-square test**

```{r}
for(col in far_cols){
  print(col)
  print(table(facility = far_data$Facility,
              far_data[[col]]))
  cat("\n\n\n")
  print(paste("Chi-squred test of", col))
  print(fisher.test(table(facility = far_data$Facility,
              far_data[[col]])))
}
```

**Interpret**

`face to face` data is the only data that has a p-value `<0.05` which indicates that the it is the best metric for predicting if individuals need the facilities

### Logistic model

considering the data set is small there is no split to train and test cases to ensure accuracy of prediction

#### Prepare the data

```{r}
far_data = far_data[!is.na(far_data$Facility), ]
```

```{r}
library(caTools)
library(ROCR)
```

```{r}
far_model = glm(Facility~.,
                    data = far_data, family = binomial)
summary(far_model)
```

**Seeing if the key predictor value improves the model prediction**

```{r}
far_model_2 = glm(Facility~ face_2_face,
                    data = far_data, family = binomial)
summary(far_model_2)
```

The second model is better at discriminating if individuals will need the facility or not as the AIC value is `1.163987%` smaller

This means that a person will be less likely to need the facilities if they are physical than if they are not physical.

### Prediction and Accuracy

```{r}
pred_far = predict(far_model_2)
pred_far = ifelse(pred_far> 0.5, 1, 0)
reg_tab_far = table( actual_data = far_data$Facility , predicted = pred_far[1:271])
reg_tab_far
```

#### accuracy

```{r}
acc_far = mean(pred_far == far_data$Facility)
acc_far
```

This model is `56.82%` accurate. This means that for every 100 individuals that are selected and need the facility. The model is able to accurately select `57` individuals while it mislabels the remaining `43` individuals as people who do not need the facility

```{r}
odds_rat_far = exp(coef(far_model_2))
odds_rat_far
```

The second model also predicts that

$$ f(need\ facility\ YES |face\ to\ face) = \frac{1}{1+ e^{-2.66667 - 0.5145\ face\ to\ face)}}  $$

This backs up the initial model coefficient claim that a person is less likely to need the facilities if they are face to face
